greta.Rmd

### 2019-10-08 notebook/appendix.Rmd
WIP

```{r}
damage <- 0.05
control <- 1
discount <- 0.98

n_s <- 200
states <- seq(0,120, length=n_s)
actions <- states

p <- list(r = .8, K = 153, q = 2, b = 20, sigma = .02, a=2.3,
          N = 3e3, x0 = 20)
#eps <- states[2]/10

may <- function(a){  
  function(x, h = 0){ # May
    x + x * p$r * (1 - x / p$K)  - a * x ^ p$q / (x ^ p$q + p$b ^ p$q) - h
  }
}

reward_fn <- function(x,h) - damage * (x / max(states)) ^ 2 - control * h / max(actions)

@@

Range of possible a that covers tipping in both directions:

@@

true_a <- 27.4      ### 21.5
believe_a <- 29 # 28.5   ### 22.5
possible_a <- seq(2*true_a - believe_a, 2*believe_a - true_a, by = 0.1)
true_i <- which.min(abs(possible_a - true_a))
```

@@

```{r}
X <- numeric(length(states))
X[x0] <- 1

@@

@@

```{r}
samples <- examples %>% group_by(series) %>% group_modify(~estimate_posterior(.x))
```

@@

```{r}
no_policy <- numeric(length(states)) + 1
x0 <- which.min(abs(states - 25))

set.seed(12345678)
df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = no_policy, x0 = x0, Tmax = Tmax)

```{r}
soln <- mdp_value_iteration(m_belief, reward, discount)
opt_soln <- mdp_value_iteration(m_true, reward, discount)
#soln <- mdplearning::mdp_compute_policy(list(m_belief), reward, discount)
#opt_soln <- mdplearning::mdp_compute_policy(list(m_true), reward, discount)
```

```{r}
tibble(state = states,
       belief = actions[soln$policy],
       ideal =  actions[opt_soln$policy]) %>%
  pivot_longer(c(belief, ideal), 
               names_to="model",
               values_to = "action") %>%
  ggplot(aes(state,action, color = model)) + geom_point() 
```

```{r}
set.seed(12345)
df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = soln$policy, x0 = x0, Tmax = 200)

df <- mdp_planning(m_belief, m$reward, discount, model_prior = c(1), 
                   policy = soln$policy, x0 = x0, Tmax = 100)

df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = opt_soln$policy, x0 = x0, Tmax = 200)

```{r}
sim <- mdp_learning(transition, reward, discount, 
                    x0 = x0, 
                    Tmax = Tmax, 
                    true_transition = transition[[true_i]],
                    model_prior = prior,
                    type = "value iteration", 
                    epsilon = 1e-2)
```

```{r}
 sim$df %>% 
  select(-value) %>% 
  gather(series, state, -time) %>% 
  ggplot(aes(time, states[state], color = series)) + geom_line()
```

```{r}
 sim$posterior %>% 
  data.frame(time = 1:Tmax) %>%
  filter(time %in% seq(1,Tmax, by = 5)) %>%
  gather(param, probability, -time, factor_key =TRUE) %>% 
  mutate(param = as.numeric(param)) %>% 
  ggplot(aes(param, probability, group = time, alpha = time)) + 
  geom_line()
```

>>>>

```{r}
#damage <- 0.3
#control <- 100
#discount <- 0.99

damage <- 0.05
control <- 1
discount <- 0.98

n_s <- 110
states <- seq(0,110, length=n_s)
actions <- states

p <- list(r = .8, K = 153, q = 2, b = 20, sigma = .02, a=2.3, x0 = 20)
eps <- states[2] / 10

may <- function(a){  
  function(x, h = 0){ # May
    pmax(
      x + x * p$r * (1 - x / p$K)  - a * x ^ p$q / (x ^ p$q + p$b ^ p$q) + eps - h,  
      0)
  }
}

reward_fn <- function(x,h) - damage * (x / 100) - control * h / 100

#reward_fn <- function(x,h) - damage * (x / max(states)) - control * h / max(actions) * x / max(states)

@@

Range of possible a that covers tipping in both directions.  Belief is in a stable system while the reality is a transient (smaller `a`).

@@

true_a <- 27.4      ### 21.5
believe_a <- 29 # 28.5   ### 22.5

possible_a <- seq(2*true_a - believe_a, 2*believe_a - true_a, by = 0.2)
true_i <- which.min(abs(possible_a - true_a))
```

@@

```{r}
X <- numeric(length(states))
X[2] <- 1

@@

```{r}
x0 <- which.min(abs(states - 23))
```

```{r}
no_policy <- numeric(length(states)) + 1

#set.seed(12345678)
set.seed(12345)

df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = no_policy, x0 = x0, Tmax = Tmax)

```{r}
belief_p <- mdp_value_iteration(m_belief, reward, discount, epsilon = 1e-4)
true_p <- mdp_value_iteration(m_true, reward, discount, epsilon = 1e-5)
#soln <- mdplearning::mdp_compute_policy(list(m_belief), reward, discount)
#opt_soln <- mdplearning::mdp_compute_policy(list(m_true), reward, discount)
```

```{r}
tibble(state = states,
       belief = actions[belief_p$policy],
       ideal =  actions[true_p$policy]) %>%
  pivot_longer(c(belief, ideal), 
               names_to="model",
               values_to = "action") %>%
  ggplot(aes(state,action, color = model)) + geom_point() 
```

```{r}
set.seed(12345)
df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = belief_p$policy, x0 = x0, Tmax = 1000)

```{r}
set.seed(12345)
df <- mdp_planning(m_belief, reward, discount, model_prior = c(1), 
                   policy = belief_p$policy, x0 = 26, Tmax = 1000)

df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = true_p$policy, x0 = x0, Tmax = 1000)

## Active Learning

```{r}
if(!file.exists("learning_sim.tsv.gz")){ # Cache the slow step

  learning_sim <- mdp_learning(transition, reward, discount, 
                      x0 = x0, 
                      Tmax = Tmax, 
                      true_transition = transition[[true_i]],
                      model_prior = prior,
                      type = "value iteration", 
                      epsilon = 1e-2)
  readr::write_tsv(learning_sim, "learning_sim.tsv.gz")

} else {
  learning_sim <- readr::read_tsv("learning_sim.tsv.gz")
}
```

```{r}
 learning_sim$df %>% 
  select(-value) %>% 
  gather(series, state, -time) %>% 
  ggplot(aes(time, states[state], color = series)) + geom_line()
```

```{r}
 learning_sim$posterior %>% 
  data.frame(time = 1:Tmax) %>%
  filter(time %in% seq(1,Tmax, by = 5)) %>%
  gather(param, probability, -time, factor_key =TRUE) %>% 
  mutate(param = as.numeric(param)) %>% 
  ggplot(aes(param, probability, group = time, alpha = time)) + 
  geom_line()
```

@@

```{r mcmc, cache=TRUE}
if(!file.exists("samples.tsv.gz")){ # Manually cache the slow step

samples <- examples %>% 
  group_by(series) %>% 
  group_modify(~estimate_posterior(.x))

readr::write_tsv(samples, "samples.tsv.gz")

} else {
  samples <- readr::read_tsv("samples.tsv.gz")
}
```

@@
